Inferring ecological interactions

### 2017-02-09 ms/ms_interactions.tex
Continued working on the method section of the manuscript

\section{Title:}\label{title}

@@

\section{Introduction}\label{introduction}

@@

In spite of increasing efforts, it is unlikely that ecologists will empirically
document the enormous number of interactions existing in nature
(Morales-Castilla et al. 2015). Therefore, complementing field work with models
of species interactions is a promising avenue for understanding the complexity
of the natural world(ref). While direct inference is often preferable, many
interactions are rare, infrequent, or impossible to detect through human
observation. Model-based inference for ecological networks provides three
important aspects to characterize species interactions, i) it establishes
mechanistic relationships among species, ii) it estimates uncertainty given
empirical observation, and iii) it provides an avenue to predict future
interactions. We follow Poisot (2015) and Cazelles (2016) in formulating our
general conceptual model of species interactions (Fig. 1). There is a pool of
true interactions, a metaweb, in which individuals of different species co-occur
resulting in interactions among species. Any given observation we make on this
metaweb is an attempt to recover the relationships among species using empirical
data. Based on these data, model(s) are built to characterize and elucidate the
relationships among species. Some models rely exclusively co-occurrence of
individuals of different species at particular locations and may also include
information on phylogenetic history and species traits. These models can than be
used to predict the interactions among observed species, given an uncertainty
due to the spatial and/or temporal stochasticity of the interactions. With these
models it is also possible to define new network, based on the presence of a
novel species in the assemblage. From these predictions, we can estimate
structural network properties, such as connectance and nestedness, while
simultaneously propagating our uncertainty throughout the analysis. In this
paper, we reviewed the methods proposed to build models of species interactions,
emphasizing each method's utility for predicting the position of a novel species
in the network.

@@

\subsection{Models of species
interactions}\label{models-of-species-interactions}

\subsubsection{Regression-based Models}\label{regression-based-models}

The simplest model of species interactions is

eq i.

\[Y_{i,j} = \rho_{i,j}\] Where the probability of interaction among species i
and species j is a fixed value (\(\rho_{i,j}\)). Using empirical data, we might
try to estimate this value by dividing the number of observations by the total
number of observations for any pair of species. While this is straightforward,
why might it be insufficient? Foremost, it contains no estimate of uncertainty.
The estimate is given without any consideration for the variance due to
sampling, spatial or temporal stochasticity. To add uncertainty, we draw our
observations from a statistical distribution.

eq ii.

\[ Y_{i,j} \sim Bernoulli(\rho_{i,j}) \]

We can now estimate the probability of interaction, as well as the uncertainty
based on sampling stochasticity. Yet this model is still unsatisfying. It is
non-mechanistic, it depends wholly on species identity, and uses no ecological
theory or inference. Therefore, it is difficult to ascribe meaning to a set of
interactions, or understand how they might change in the future. We cannot use
the inferences from data to make a prediction about a new species not in
dataset.

To make our model more useful, we might add a covariate, such that

eq. iii

\[ Y_{i,j} \sim Bernoulli(\rho_{i,j})\] \[ logit(\rho_{i,j}) = \alpha_{i,j} +
\beta_{i,j} * x_{i,j} \]

In which a link between species i and j is a Bernoulli trial with a probability
of success \(\rho_{i,j}\). The probability of success will vary based on some
ecological principle (\(x_{i,j}\)), for example, the similarity in functional
traits among species i and j:

eq. V

\[x_{i,j}= f(Trait_i,Trait_j)\]

@@

\subsubsection{Logic-based methods}\label{logic-based-methods}

@@

\subsubsection{Occurrence-based methods}\label{occurrence-based-methods}

@@

\subsubsection{Non-parametric methods}\label{non-parametric-methods}

The above methods are model based, that is, they largely use parametric
distributions to estimate the liklihood of the data given estimated model
parameters. An alternative approach is to use multivariate statistics to create
association matrices. One such method is called fourth-corner analysis, which is
a type of three table ordination (sometimes called `RLQ' analysis) (Doledec
1996). In bipartite networks, fourth-corner analysis relates the matrix of
species interactions to a matrix of traits from one level (e.g.~plants) to
another level (e.g.~pollinators) via the matrix of species innteractions (Dray \&
Legendre, 2008). This produces a correlation matrix, which is then compared to a
null expectation through matrix randomization. Randomization is a familiar tool
in network ecology, in which permuatation rules are used to compare the observed
test statistic (e.g the correlation among species occurrences) to a null
distribution based on some matrix constraint. Common constraints include the
mantainance of row sums, column sums, or other all marginal values, such that
only the identity of partners are switched. After each permutation, the test
statistic is recalulcating, yielding a null distribution. By comparing whether
the observed test statistic falls within the upper or lower 5th quantile of the
null distribution, studies distinguish whether their results could have been
generated at random given their sampling effort.

\section{Methods}\label{methods}

Our goal is provide a conceptual framework for the broad array of potential
methods described above. Our three criteria for qualitative comparison are: 1)
Can the method account for the uncertainty in species observations?, 2) Does it
provide an ecological mechanism to infer the probability of species
interactions, and 3) Can it be used to predict an novel interaction? In addition
to qualitative comparison, we provide a quantitative analysis of methods for
inferring species interactions. From a quantitative perspective, we are
interested in methods which accurately reflect true relationships among data,
are relatively insensitive to sampling effort, and make reasonable assumptions
about the interdependence of species observations. We use a combination of
simulated and empirical data to establish a benchmark dataset to compare the
accuracy and sensitivity of currently proposed methods.

\section{Quantitative comparisons}\label{quantitative-comparisons}

@@

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\subsection{Perspectives}\label{perspectives}

@@

\subsection{On the virtue of prediction of species
interactions.}\label{on-the-virtue-of-prediction-of-species-interactions.}

@@

\section{Literature cited}\label{literature-cited}

@@

\section{Boxes}\label{potential-boxes}

\subsection{Soil ecology}\label{soil-ecology}

@@

\subsection{Climate changes and alien
species}\label{climate-changes-and-alien-species}

\subsection{Community/evolution ecology}\label{communityevolution-ecology}

>>>>

\section{Title:}
\label{title}

@@

\section*{Abstract}
\label{abstract}

\section*{Introduction}
\label{introduction}

@@

In spite of increasing efforts, it is unlikely that ecologists will empirically
document the enormous number of interactions existing in nature
(Morales-Castilla et al. 2015). Therefore, complementing fieldwork with models
of species interactions is a promising avenue for understanding the complexity
of the natural world (ref). While direct inference is often preferable, many
interactions are rare, infrequent, or impossible to detect through human
observation. Model-based inference for ecological networks provides three
important aspects to characterize species interactions, 1) it establishes
mechanistic relationships among species, 2) it estimates uncertainty given
empirical observation, and 3) it provides an avenue to predict future
interactions. We follow Poisot (2015) and Cazelles (2016) in formulating our
general conceptual model of species interactions (Figure \ref{fig:concept}).
There is a pool of true interactions, a metaweb, in which individuals of
different species co-occur resulting in interactions among species. Any given
observation we make on this metaweb is an attempt to recover the relationships
among species using empirical data. Based on these data, model(s) are built to
characterize and elucidate the relationships among species. Some models rely
exclusively co-occurrence of individuals of different species at particular
locations and may also include information on phylogenetic history and species
traits. These models can than be used to predict the interactions among observed
species, given an uncertainty due to the spatial and/or temporal stochasticity
of the interactions. With these models it is also possible to define new
networks, based on the presence of new species in the assemblage. From these
predictions, we can estimate structural network properties such as connectance
and nestedness, while simultaneously propagating our uncertainty throughout the
analysis. In this paper, we reviewed the methods proposed to build models of
species interactions, emphasizing each method's utility for predicting the
position of a novel species in the network.

@@

\section*{Methods}\label{methods}

One goal of this paper is to frame the different techniques used to model
ecological networks within a conceptual framework that would allow ecologists to
efficiently choose which technique to use with regard to various properties of
the methods and the type of ecological questions they can answer. A way to
assess the different properties of the modelling techniques is to compare them
based on specific criteria. We defined three criteria, formulated as questions,
that can be used to qualitatively compare the methods used to model ecological
networks: 1) Can the method account for the uncertainty in species
observations?; 2) Does the method provide ecological mechanisms to infer the
probability of species interactions?; and 3) Can the method be used to predict
an novel interaction?

In addition to qualitative comparison, we also provide a quantitative comparison
of methods for inferring species interactions. From a quantitative perspective,
we are interested in methods that: 1) accurately reflect the true underlying
relationships among species, 2) are insensitive to sampling effort, and 3) make
reasonable assumptions about the interdependence of species observations. To
quantitatively compare the different methods, we used simulated and reference
empirical data to establish benchmarks that will later be used to compare the
accuracy and sensitivity of currently proposed methods.

\subsection*{Modelling ecological networks - A diversity of methods}
\label{models-of-species-interactions}

When modelling ecological networks, there are different types of methods that
have been proposed in the litterature. In this section, we presented the
different techniques and group them based on their properties.

\subsubsection*{Co-occurrences models}
\label{regression-based-models}
%%% Regression-based models on co-occurrences subsubsection will probably have to be moved somewhere else to follow the structure of the Introduction where traits and phylogeny related methods are discussed firsts and co-occurrence methods are discussed later.

The simplest model of that can be designed to characterize the co-occurrence
patterns between species $i$ and $j$ in matrix $Y$ can be defined as follow

\begin{equation}
  \label{eq:coOccSimple}
  Y_{i,j} = \rho_{i,j}
\end{equation}

where the probability of interaction among species $i$ and $j$ is a fixed value
($\rho_{i,j}$). Using empirical data, we might try to estimate this value by
dividing the number of observations by the total number of observations for any
pair of species. While this is straightforward, it is insufficient because it
contains no estimate of uncertainty. The estimate is given without any
consideration for the variance due to sampling, environmental, spatial or
temporal stochasticity. This model is overly simple. By assuming that
observations are randomly drawn from a statistical distribution with parameter
$\rho_{i,j}$, we obtain a simple model that account for sampling stochasticity
in the data.

\begin{equation}
  \label{eq:coOccBern}
  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
\end{equation}

In equation \ref{eq:coOccBern}, we used a Bernoulli distribution, which assumes
that $Y_{i,j}$ is either 0 or 1. If $Y_{i,j}$ includes other types of data (e.g.
counts or weights) other distributions may be used instead. From equation
\ref{eq:coOccBern}, we can now estimate the probability of interaction while
accounting sampling stochasticity. Although this model is more informative than
the one presented in equation \ref{eq:coOccSimple}, it is still unsatisfying
because it depends wholly on species identity and uses no ecological theory or
inference; in short it is not mechanistic. Therefore, it is difficult to ascribe
meaning to the species association that could be highlighted by this model, or
understand how they might change in the future. In addition, from this model, we
cannot make a prediction about potential new species.

To make our model more ecologically relevant, we might add covariates $X_{i,j}$
that influence the relationship among a pair of species

\begin{equation}
  \label{eq:coOccCov}
  Y_{i,j} \sim \text{Bernoulli}(\rho_{i,j})
  \text{logit}(\rho_{i,j}) = \alpha_{i,j}+\beta_{i,j}X_{i,j}
\end{equation}

where $\beta_{i,j}$ is a regression coefficient and $\alpha_{i,j}$ is the model
intercept. In equation \ref{eq:coOccCov}, a link between species $i$ and $j$ is
a Bernoulli trial with a probability of success $\rho_{i,j}$. The probability of
success will vary based on some ecological principle ($x_{i,j}$) such as the
similarity in functional traits among species $i$ and $j$:

%%% GB: I am not sure if this equation is right... Need to check carefully.
\begin{equation}
  \label{eq:coOccTrait}
  x_{i,j}= f(\text{Trait}_i,\text{Trait}_j)
\end{equation}

%%% ICI %%%

@@

\subsubsection*{Logic-based methods}\label{logic-based-methods}

@@

\subsubsection*{Occurrence-based methods}\label{occurrence-based-methods}

@@

\subsubsection*{Non-parametric methods}\label{non-parametric-methods}

The above methods are model based, that is, they largely use parametric
distributions to estimate the likelihood of the data given estimated model
parameters. An alternative approach is to use multivariate statistics to create
association matrices. One such method is called fourth-corner analysis, which is
a type of three table ordination (sometimes called `RLQ' analysis) (Doledec
1996). In bipartite networks, fourth-corner analysis relates the matrix of
species interactions to a matrix of traits from one level (e.g.~plants) to
another level (e.g.~pollinators) via the matrix of species innteractions (Dray \&
Legendre, 2008). This produces a correlation matrix, which is then compared to a
null expectation through matrix randomization. Randomization is a familiar tool
in network ecology, in which permuatation rules are used to compare the observed
test statistic (e.g the correlation among species occurrences) to a null
distribution based on some matrix constraint. Common constraints include the
mantainance of row sums, column sums, or other all marginal values, such that
only the identity of partners are switched. After each permutation, the test
statistic is recalulcating, yielding a null distribution. By comparing whether
the observed test statistic falls within the upper or lower 5th quantile of the
null distribution, studies distinguish whether their results could have been
generated at random given their sampling effort.

\subsection*{Quantitative comparisons}\label{quantitative-comparisons}

@@

\section*{Results}\label{results}

\section*{Discussion}\label{discussion}

\subsection*{Perspectives}\label{perspectives}

@@

\subsection*{On the virtue of prediction of species
interactions.}\label{on-the-virtue-of-prediction-of-species-interactions.}

@@

\section*{Literature cited}\label{literature-cited}

@@

\section*{Boxes}\label{potential-boxes}

\subsection*{Soil ecology}\label{soil-ecology}

@@

\subsection*{Climate changes and alien species}
\label{climate-changes-and-alien-species}

\subsection*{Community/evolution ecology}
\label{communityevolution-ecology}
